An introduction to wavelets using R

### 2024-07-07 text/24/wavelets.Rmd
Updated 2024 Rmd files

---
title: "An introduction to wavelets using R"
author: "Steve Simon"
source: "http://www.pmean.com/wavelets.html"
date: "2014-01-01"
category: 
- Blog post
tags:
- Checking assumptions  
- Linear regression
- R software
output: html_document
page_update: partial
---

@@

```{r setup}
knitr::opts_chunk$set(fig.height=1)
```

```{r read-ecg-data}
library("ggplot2")
library("magrittr")
library("rwt")
library("wavelets")
data(ecg)
head(ecg)
tail(ecg)
```

@@

```{r graphing-functions}
graph_with_axis <- function(x, lab, limits=range(x)) {
  par(mar=c(2.6, 0, 0, 0))
  plot((1:2048)/2048, x, 
       xlim=c(0, 1.2), 
       ylim=limits, 
       type="l",
       axes=FALSE)
  axis(side=1, at=(0:4)/4, labels=c(1, 512, 1024, 1536, 2048))
  text(1.1, x[2048], lab, adj=0.5, cex=0.67)
}
graph_without_axis <- function(x, lab, limits=range(x)) {
  par(mar=c(0, 0, 0, 0))
  plot((1:2048)/2048, x, 
       xlim=c(0, 1.2), 
       ylim=c(limits[1], limits[2]),
       type="l",
       axes=FALSE)
  text(1.1, x[2048], lab, adj=0.5, cex=0.67)
}
needle_plot <- function(y, lab, x=seq(0, 1, along.with=y), limits=range(y)) {
  par(mar=c(0, 0, 0, 0))
  plot(
    x, y, 
    xlim=c(0, 1.2), 
    ylim=limits,
    type="n",
    axes=FALSE)
  segments(x, 0, x, y)
  text(1.1, 0, lab, cex=0.67)
}
```

@@

```{r graph-ecg, fig.height=2}
graph_with_axis(ecg, "ecg\nsignal")
```

@@

```{r apply-fft}
ft.ecg <- fft(ecg)/2048
head(ft.ecg)
tail(ft.ecg)
```

@@

```{r complex-plot, fig.height=5}
par(mar=c(0,0,0,0))
plot(Re(ft.ecg), Im(ft.ecg), axes=FALSE, type="n")
text(Re(ft.ecg), Im(ft.ecg), 1:2048)
```

@@

```{r inverse-fft}
ift.ecg <- fft(ft.ecg,inverse=TRUE)
head(ift.ecg)
tail(ift.ecg)
```

@@

```{r compute-mean}
mean(ecg)
ft.ecg[1]
```

@@

```{r graph-cosine-1}
t.unit <- (1:2048)/2048
cos1 <- +Re(ft.ecg[2])*cos(2*pi*t.unit)
sin1 <- -Im(ft.ecg[2])*sin(2*pi*t.unit)
sum1 <- cos1+sin1

@@

```{r compare-to-original-curve}
{
  graph_without_axis(sum1, "Combined\ncomponent")
  graph_without_axis(ecg, "Original\nsignal")
}
```

@@

```{r second-component}
ft.comp2 <- ft.ecg
ft.comp2[-2] <- 0
comp2 <- Re(fft(ft.comp2,inverse=TRUE))
graph_without_axis(sum1, "Using\nInverse fft")
{
  lines((1:2048)/2048, sum1, col="red", lty="dashed", lwd=3)
}
```

@@

```{r graph-third-component}
ft.comp3 <- ft.ecg
ft.comp3[-3] <- 0
comp3 <- Re(fft(ft.comp3,inverse=TRUE))
graph_without_axis(comp3, "Third\ncomponent")
```

@@

```{r graph-fourth-component}
ft.comp4 <- ft.ecg
ft.comp4[-4] <- 0
comp4 <- Re(fft(ft.comp4,inverse=TRUE))
graph_without_axis(comp4, "Fourth\ncomponent")
```

@@

```{r thirty-second}
ft.comp32 <- ft.ecg
ft.comp32[-32] <- 0
comp32 <- Re(fft(ft.comp32,inverse=TRUE))
graph_without_axis(comp32, "32nd\ncomponent")
```

@@

```{r count}
m <- max(comp32)
graph_without_axis(comp32, "32nd\ncomponent", limits=c(-m, 1.2*m))
text((seq(1, 31, by=3)-0.5)/31, 1.2*m, seq(1, 31, by=3), col="red", cex=0.67) 
```

@@

```{r rwt-package}
d4.rwt <- daubcqf(4)
d4.rwt
wt.rwt <- mdwt(ecg,daubcqf(4)$h.0,L=9)
names(wt.rwt)
head(wt.rwt$y)
tail(wt.rwt$y)
wt.rwt$L
iwt.rwt <- midwt(wt.rwt$y,daubcqf(4)$h.0,L=9)
names(iwt.rwt)
head(iwt.rwt$x)
tail(iwt.rwt$x)
iwt.rwt$L
```

@@

```{r wavelets-package}
d4.wavelets <- wt.filter("d4")
d4.wavelets
wt.wavelets <- dwt(ecg,filter="d4")
slotNames(wt.wavelets)
iwt.wavelets <- idwt(wt.wavelets)
head(iwt.wavelets)
tail(iwt.wavelets)
```

@@

```{r graph-d4}
wt.comp <- function(x, y) {
  single_wavelet <- as.numeric((1:2048)==y)
  as.vector(midwt(single_wavelet, daubcqf(4)$h.0, L=8)$x)
}
```

```{r calculate-functions}
calculate_mother <- function(i, L) {
  yl <- rep(0, 2048)
  yh <- matrix(0, nrow=2048, ncol=L)
  yh[i, L] <- 1
  as.vector(mirdwt(yl, yh, daubcqf(4)$h.0, L)$x)
}
calculate_father <- function(i, L) {
  yl <- rep(0, 2048)
  yh <- matrix(0, nrow=2048, ncol=L)
  yl[i] <- 1
  as.vector(mirdwt(yl, yh, daubcqf(4)$h.0, L)$x)
}
```

```{r mother-and-father}
{
  graph_without_axis(calculate_father(2^9+1, 7),"d4 father")
  graph_without_axis(calculate_mother(2^9+1, 7),"d4 mother")
}
```

@@

```{r shifted}
{
  graph_without_axis(calculate_father(2^9-2^7+1, 7),"d4 father\nshifted left")
  graph_without_axis(calculate_father(2^9+2^7+1, 7),"d4 father\nshifted right")
  graph_without_axis(calculate_mother(2^9-2^7+1, 7),"d4 mother\nshifted left")
  graph_without_axis(calculate_mother(2^9+2^7+1, 7),"d4 mother\nshifted right")
}
```

@@

```{r wrapped}
{
  graph_without_axis(calculate_father(2^10+2^9+1+2^8+1, 7),"d4 father, wrapped")
  graph_without_axis(calculate_mother(2^10+2^9+1+2^8+1, 7),"d4 mother, wrapped")
}

@@

```{r stretch-and-squeeze}
{  
  graph_without_axis(calculate_father(2^9+1, 8),"d4 father\nstretched")
  graph_without_axis(calculate_father(2^9+1, 6),"d4 father\nsqueezed")
  graph_without_axis(calculate_mother(2^9+1, 8),"d4 mother\nstretched")
  graph_without_axis(calculate_mother(2^9+1, 6),"d4 mother\nsqueezed")
}
```

@@

```{r most-extreme}
{
  graph_without_axis(calculate_father(2^9+1, 1),"d4 father\nmost extreme")
  graph_without_axis(calculate_mother(2^9+1, 1),"d4 mother\nmost extreme")
}
```

@@

```{r high-pass}
g <- daubcqf(4)$h.0
g
```

@@

```{r low-pass}
h <- -daubcqf(4)$h.1
h
```

@@

```{r orthonormal}
# The inline function %*% calculates a dot product of two vectors.
g %*% g
h %*% h
g %*% h
```

@@

```{r overlap}
g[1:2] %*% g[3:4]
h[1:2] %*% h[3:4]
g[1:2] %*% h[3:4]
h[1:2] %*% g[3:4]
```

@@

```{r orthogonal-to-linear}
h %*% (1:4)
```

@@

```{r first-level-wavelet}
wt.1 <- mdwt(ecg, daubcqf(4)$h.0, L=1)
V1 <- wt.1$y[   1:1024]
W1 <- wt.1$y[1025:2048]
{
  needle_plot(V1, "V1", limits=range(ecg))
  needle_plot(W1, "W1", limits=range(ecg))
  graph_without_axis(ecg, "Original\nsignal", limits=range(ecg))
}
```

@@

```{r first-level-dot-products}
c(V1[   1], g %*% ecg[1:4])
c(V1[   2], g %*% ecg[3:6])
c(V1[   3], g %*% ecg[5:8])
c(V1[1023], g %*% ecg[2045:2048])
c(V1[1024], g %*% ecg[c(2047:2048, 1:2)])

@@

```{r second-level-wavelet}
wt.2 <- mdwt(ecg, daubcqf(4)$h.0, L=2)
V2 <- wt.2$y[1:512]
W2 <- wt.2$y[513:1024]
quality_check_W1 <- wt.2$y[1025:2048]
all(quality_check_W1==W1)
{
  needle_plot(V2, "V2", limits=range(ecg))
  needle_plot(W2, "W2", limits=range(ecg))
  needle_plot(W1, "W1", limits=range(ecg))
  graph_without_axis(ecg, "Original\nsignal", limits=range(ecg))
}
```

```{r second-level-dot-products}
c(V2[  1], g %*% V1[1:4])
c(V2[  2], g %*% V1[3:6])
c(V2[  3], g %*% V1[5:8])
c(V2[511], g %*% V1[1021:1024])
c(V2[512], g %*% V1[c(1023:1024, 1:2)])

@@

```{r transform-V1}
quality_check <- mdwt(V1, daubcqf(4)$h.0, L=1)
quality_check_V2 <- quality_check$y[  1: 512]
quality_check_W2 <- quality_check$y[513:1024]
all(quality_check_V2==V2)
all(quality_check_W2==W2)
```

@@

```{r third-level-wavelet}
wt.3 <- mdwt(ecg, daubcqf(4)$h.0, L=3)
V3 <- wt.3$y[1:256]
W3 <- wt.3$y[257:512]
W2_check3 <- wt.3$y[513:1024]
W1_check3 <- wt.3$y[1025:2048]
all(W2_check3==W2)
all(W1_check3==W1)
{
  needle_plot(V3, "V3", limits=range(ecg))
  needle_plot(W3, "W3", limits=range(ecg))
  needle_plot(W2, "W2", limits=range(ecg))
  needle_plot(W1, "W1", limits=range(ecg))
  graph_without_axis(ecg, "Original\nsignal", limits=range(ecg))
}
```

@@

```{r third-level-dot-products}
c(V3[  1], g %*% V2[1:4])
c(V3[  2], g %*% V2[3:6])
c(V3[  3], g %*% V2[5:8])
c(V3[255], g %*% V2[509:512])
c(V3[256], g %*% V2[c(511:512, 1:2)])

@@

```{r ninth-level-wavelet}
wt.9 <- mdwt(ecg, daubcqf(4)$h.0, L=9)
V9 <- wt.9$y[   1:   4]
W9 <- wt.9$y[   5:   8]
W8 <- wt.9$y[   9:  16]
W7 <- wt.9$y[  17:  32]
W6 <- wt.9$y[  33:  64]
W5 <- wt.9$y[  65: 128]
W4 <- wt.9$y[ 129: 256]
W3 <- wt.9$y[ 257: 512]
W2 <- wt.9$y[ 513:1024]
W1 <- wt.9$y[1025:2048]
{
  needle_plot(V9, "V9", limits=range(ecg))
  needle_plot(W9, "W9", limits=range(ecg))
  needle_plot(W8, "W8", limits=range(ecg))
  needle_plot(W7, "W7", limits=range(ecg))
  needle_plot(W6, "W6", limits=range(ecg))
  needle_plot(W5, "W5", limits=range(ecg))
  needle_plot(W4, "W4", limits=range(ecg))
  needle_plot(W3, "W3", limits=range(ecg))
  needle_plot(W2, "W2", limits=range(ecg))
  needle_plot(W1, "W1", limits=range(ecg))
  graph_without_axis(ecg, "Original\nsignal", limits=range(ecg))
}
```

@@

```{r reconstuct-V8}
V9_diluted <- rep(0, 8)
V9_diluted[(1:4)*2-1] <- V9
W9_diluted <- rep(0, 8)
W9_diluted[(1:4)*2-1] <- W9

@@

```{r reconstuct-V7}
V8_diluted <- rep(0, 16)
V8_diluted[(1:8)*2-1] <- V8
W8_diluted <- rep(0, 16)
W8_diluted[(1:8)*2-1] <- W8

@@

```{r reconstuct-V6}
V7_diluted <- rep(0, 32)
V7_diluted[(1:16)*2-1] <- V7
W7_diluted <- rep(0, 32)
W7_diluted[(1:16)*2-1] <- W7

@@

```{r graph-v}

@@

```{r graph-w}

@@

```{r read-dog-image}
knitr::opts_chunk$set(fig.width=5, fig.height=5)
fn <- "https://www2.isye.gatech.edu/isyebayes/data/klaus256.asc"
img <- read.table(fn)
img <- 2*img-1
```

```{r display-dog-image}
img.plot <- function(imat,n=256) {
  xsq <- c(0,0,1,1)
  ysq <- c(0,1,1,0)
  par(mar=rep(0,4),xaxs="i",yaxs="i")
  plot(c(1,n+1),c(1,n+1),type="n",axes=FALSE)
  # imat <- imat/max(abs(imat))
  for (i in 1:n) {
    for (j in 1:n){
      d <- 0.5*imat[n+1-j,i]+0.5
      d <- min(1,max(0,d))
      polygon(xsq+i,ysq+j,density=-1,border=NA,col=rgb(d,d,d))
    }
  }
}

@@

```{r d8-wavelet-g}
g <- daubcqf(8)$h.0
g
```

```{r d8-wavelet-h}
h <- -daubcqf(8)$h.1
h
```

@@

```{r d8-in-1d, fig.height=1}
d8_father_coefficients <- rep(0, 2048)
d8_father_coefficients[3] <- 1
d8_father <- as.vector(midwt(d8_father_coefficients, daubcqf(8)$h.0, L=8)$x)
d8_mother_coefficients <- rep(0, 2048)
d8_mother_coefficients[10] <- 1
d8_mother <- as.vector(midwt(d8_mother_coefficients, daubcqf(8)$h.0, L=8)$x)
{
  graph_without_axis(d8_father, "D8 father")
  graph_without_axis(d8_mother, "D8 mother")
}
```

@@

```{r decompose-and-restructure}
wt.img.4 <- mdwt(img, daubcqf(8)$h.0, L=4)
reconstructed.img <- midwt(wt.img.4$y, daubcqf(8)$h.0, L=4)
img.plot(reconstructed.img$x)
```

```{r check-2d-calculations}
wt.img.1 <- mdwt(img, daubcqf(8)$h.0, L=1)
c(sum(img[1:8,1:8]*outer(g,g)), wt.img.1$y[ 1,   1])
c(sum(img[1:8,1:8]*outer(g,h)),wt.img.1$y[  1, 129])
c(sum(img[1:8,1:8]*outer(h,g)),wt.img.1$y[129,   1])
c(sum(img[1:8,1:8]*outer(h,h)),wt.img.1$y[129, 129])
```

@@

```{r layout}
x <- c(700, 300, 700)
y <- c(700, 300, 300)
m <- c(
  "W1\nh %*% g", 
  "W1\ng %*% h", 
  "W1\nh %*% h"
)
data.frame(x, y, m) %>%
  ggplot(aes(x, y, label=m)) +
  expand_limits(x=c(100, 900), y=c(100, 900)) + 
  geom_segment(x=100, y=100, xend=900, yend=100) +
  geom_segment(x=900, y=100, xend=900, yend=900) +
  geom_segment(x=900, y=900, xend=100, yend=900) +
  geom_segment(x=100, y=900, xend=100, yend=100) +
  geom_segment(x=500, y=100, xend=500, yend=900) +
  geom_segment(x=100, y=500, xend=900, yend=500) +
  theme_void() + 
  geom_text() -> layout_1
layout_1 +
    geom_text(x=300, y=700, label="V1\ng %*% g")
layout_1+
  geom_segment(x=300, y=500, xend=300, yend=900) +
  geom_segment(x=100, y=700, xend=500, yend=700) +
  geom_text(x=400, y=800, label="W2\nh %*% g") + 
  geom_text(x=200, y=600, label="W2\ng %*% h") + 
  geom_text(x=400, y=600, label="W2\nh %*% h") -> layout_2
layout_2 +
  geom_text(x=200, y=800, label="V2\ng %*% g")
layout_2 +
  geom_segment(x=200, y=700, xend=200, yend=900) +
  geom_segment(x=100, y=800, xend=300, yend=800) +
  geom_text(x=250, y=850, label="W3") + 
  geom_text(x=150, y=750, label="W3") + 
  geom_text(x=250, y=750, label="W3") + 
  geom_text(x=150, y=850, label="etc.")

@@

```{r level-1-image-g-g}
gg1 <- matrix(0, nrow=256, ncol=256)
gg1[1:128, 1:128] <- wt.img.1$y[1:128, 1:128]
gg1_signal <- midwt(gg1, daubcqf(8)$h.0, L=1)  
img.plot(gg1_signal$x)
```

```{r level-1-image-g-h}
gh1 <- matrix(0, nrow=256, ncol=256)
gh1[1:128, 129:256] <- wt.img.1$y[1:128, 129:256]
gh1_signal <- midwt(gh1, daubcqf(8)$h.0, L=1)  
img.plot(gh1_signal$x)
```

```{r level-1-image-h-g}
hg1 <- matrix(0, nrow=256, ncol=256)
hg1[1:128, 129:256] <- wt.img.1$y[129:256, 1:128]
hg1_signal <- midwt(hg1, daubcqf(8)$h.0, L=1)  
img.plot(hg1_signal$x)
```

```{r level-1-image-h-h}
hh1 <- matrix(0, nrow=256, ncol=256)
hh1[129:256, 129:256] <- wt.img.1$y[129:256, 129:256]
hh1_signal <- midwt(hh1, daubcqf(8)$h.1, L=1)  
img.plot(hh1_signal$x)
```

>>>>

---
title: "An introduction to wavelets using R"
author: "Steve Simon"
source: "http://www.pmean.com/wavelets.html"
date: "2014-01-01"
category: 
- Blog post
tags:
- Checking assumptions  
- Linear regression
- R software
- Rmarkdown graphs
output: html_document
page_update: partial
---

@@

```{r wavelets-01}
knitr::opts_chunk$set(fig.height=1)
```

```{r wavelets-02}
library("ggplot2")
library("magrittr")
library("rwt")
library("wavelets")
data(ecg)
head(ecg)
tail(ecg)
```

@@

```{r wavelets-03}
graph_with_axis <- function(x, lab, limits=range(x)) {
  par(mar=c(2.6, 0, 0, 0))
  plot((1:2048)/2048, x, 
       xlim=c(0, 1.2), 
       ylim=limits, 
       type="l",
       axes=FALSE)
  axis(side=1, at=(0:4)/4, labels=c(1, 512, 1024, 1536, 2048))
  text(1.1, x[2048], lab, adj=0.5, cex=0.67)
}
graph_without_axis <- function(x, lab, limits=range(x)) {
  par(mar=c(0, 0, 0, 0))
  plot((1:2048)/2048, x, 
       xlim=c(0, 1.2), 
       ylim=c(limits[1], limits[2]),
       type="l",
       axes=FALSE)
  text(1.1, x[2048], lab, adj=0.5, cex=0.67)
}
needle_plot <- function(y, lab, x=seq(0, 1, along.with=y), limits=range(y)) {
  par(mar=c(0, 0, 0, 0))
  plot(
    x, y, 
    xlim=c(0, 1.2), 
    ylim=limits,
    type="n",
    axes=FALSE)
  segments(x, 0, x, y)
  text(1.1, 0, lab, cex=0.67)
}
```

@@

```{r wavelets-04, fig.height=2}
graph_with_axis(ecg, "ecg\nsignal")
```

@@

```{r wavelets-05}
ft.ecg <- fft(ecg)/2048
head(ft.ecg)
tail(ft.ecg)
```

@@

```{r wavelets-06, fig.height=5}
par(mar=c(0,0,0,0))
plot(Re(ft.ecg), Im(ft.ecg), axes=FALSE, type="n")
text(Re(ft.ecg), Im(ft.ecg), 1:2048)
```

@@

```{r wavelets-07}
ift.ecg <- fft(ft.ecg,inverse=TRUE)
ead(ift.ecg)
tail(ift.ecg)
```

@@

```{r wavelets-08}
mean(ecg)
ft.ecg[1]
```

@@

```{r wavelets-09}
t.unit <- (1:2048)/2048
cos1 <- +Re(ft.ecg[2])*cos(2*pi*t.unit)
sin1 <- -Im(ft.ecg[2])*sin(2*pi*t.unit)
sum1 <- cos1+sin1

@@

```{r wavelets-10}
{
  graph_without_axis(sum1, "Combined\ncomponent")
  graph_without_axis(ecg, "Original\nsignal")
}
```

@@

```{r wavelets-11}
ft.comp2 <- ft.ecg
ft.comp2[-2] <- 0
comp2 <- Re(fft(ft.comp2,inverse=TRUE))
graph_without_axis(sum1, "Using\nInverse fft")
{
  lines((1:2048)/2048, sum1, col="red", lty="dashed", lwd=3)
}
```

@@

```{r wavelets-12}
ft.comp3 <- ft.ecg
ft.comp3[-3] <- 0
comp3 <- Re(fft(ft.comp3,inverse=TRUE))
graph_without_axis(comp3, "Third\ncomponent")
```

@@

```{r wavelets-13}
ft.comp4 <- ft.ecg
ft.comp4[-4] <- 0
comp4 <- Re(fft(ft.comp4,inverse=TRUE))
graph_without_axis(comp4, "Fourth\ncomponent")
```

@@

```{r wavelets-14}
ft.comp32 <- ft.ecg
ft.comp32[-32] <- 0
comp32 <- Re(fft(ft.comp32,inverse=TRUE))
graph_without_axis(comp32, "32nd\ncomponent")
```

@@

```{r wavelets-15}
m <- max(comp32)
graph_without_axis(comp32, "32nd\ncomponent", limits=c(-m, 1.2*m))
text((seq(1, 31, by=3)-0.5)/31, 1.2*m, seq(1, 31, by=3), col="red", cex=0.67) 
```

@@

```{r wavelets-16}
d4.rwt <- daubcqf(4)
d4.rwt
wt.rwt <- mdwt(ecg,daubcqf(4)$h.0,L=9)
names(wt.rwt)
head(wt.rwt$y)
tail(wt.rwt$y)
wt.rwt$L
iwt.rwt <- midwt(wt.rwt$y,daubcqf(4)$h.0,L=9)
names(iwt.rwt)
head(iwt.rwt$x)
tail(iwt.rwt$x)
iwt.rwt$L
```

@@

```{r wavelets-17}
d4.wavelets <- wt.filter("d4")
d4.wavelets
wt.wavelets <- dwt(ecg,filter="d4")
slotNames(wt.wavelets)
iwt.wavelets <- idwt(wt.wavelets)
head(iwt.wavelets)
tail(iwt.wavelets)
```

@@

```{r wavelets-18}
wt.comp <- function(x, y) {
  single_wavelet <- as.numeric((1:2048)==y)
  as.vector(midwt(single_wavelet, daubcqf(4)$h.0, L=8)$x)
}
```

```{r wavelets-19}
calculate_mother <- function(i, L) {
  yl <- rep(0, 2048)
  yh <- matrix(0, nrow=2048, ncol=L)
  yh[i, L] <- 1
  as.vector(mirdwt(yl, yh, daubcqf(4)$h.0, L)$x)
}
calculate_father <- function(i, L) {
  yl <- rep(0, 2048)
  yh <- matrix(0, nrow=2048, ncol=L)
  yl[i] <- 1
  as.vector(mirdwt(yl, yh, daubcqf(4)$h.0, L)$x)
}
```

```{r wavelets-20}
{
  graph_without_axis(calculate_father(2^9+1, 7),"d4 father")
  graph_without_axis(calculate_mother(2^9+1, 7),"d4 mother")
}
```

@@

```{r wavelets-21}
{
  graph_without_axis(calculate_father(2^9-2^7+1, 7),"d4 father\nshifted left")
  graph_without_axis(calculate_father(2^9+2^7+1, 7),"d4 father\nshifted right")
  graph_without_axis(calculate_mother(2^9-2^7+1, 7),"d4 mother\nshifted left")
  graph_without_axis(calculate_mother(2^9+2^7+1, 7),"d4 mother\nshifted right")
}
```

@@

```{r wavelets-22}
{
  graph_without_axis(calculate_father(2^10+2^9+1+2^8+1, 7),"d4 father, wrapped")
  graph_without_axis(calculate_mother(2^10+2^9+1+2^8+1, 7),"d4 mother, wrapped")
}

@@

```{r wavelets-23}
{  
  graph_without_axis(calculate_father(2^9+1, 8),"d4 father\nstretched")
  graph_without_axis(calculate_father(2^9+1, 6),"d4 father\nsqueezed")
  graph_without_axis(calculate_mother(2^9+1, 8),"d4 mother\nstretched")
  graph_without_axis(calculate_mother(2^9+1, 6),"d4 mother\nsqueezed")
}
```

@@

```{r wavelets-24}
{
  graph_without_axis(calculate_father(2^9+1, 1),"d4 father\nmost extreme")
  graph_without_axis(calculate_mother(2^9+1, 1),"d4 mother\nmost extreme")
}
```

@@

```{r wavelets-25}
g <- daubcqf(4)$h.0
g
```

@@

```{r wavelets-26}
h <- -daubcqf(4)$h.1
h
```

@@

```{r wavelets-27}
# The inline function %*% calculates a dot product of two vectors.
g %*% g
h %*% h
g %*% h
```

@@

```{r wavelets-28}
g[1:2] %*% g[3:4]
h[1:2] %*% h[3:4]
g[1:2] %*% h[3:4]
h[1:2] %*% g[3:4]
```

@@

```{r wavelets-29}
h %*% (1:4)
```

@@

```{r wavelets-30}
wt.1 <- mdwt(ecg, daubcqf(4)$h.0, L=1)
V1 <- wt.1$y[   1:1024]
W1 <- wt.1$y[1025:2048]
{
  needle_plot(V1, "V1", limits=range(ecg))
  needle_plot(W1, "W1", limits=range(ecg))
  graph_without_axis(ecg, "Original\nsignal", limits=range(ecg))
}
```

@@

```{r wavelets-31}
c(V1[   1], g %*% ecg[1:4])
c(V1[   2], g %*% ecg[3:6])
c(V1[   3], g %*% ecg[5:8])
c(V1[1023], g %*% ecg[2045:2048])
c(V1[1024], g %*% ecg[c(2047:2048, 1:2)])

@@

```{r wavelets-32}
wt.2 <- mdwt(ecg, daubcqf(4)$h.0, L=2)
V2 <- wt.2$y[1:512]
W2 <- wt.2$y[513:1024]
quality_check_W1 <- wt.2$y[1025:2048]
all(quality_check_W1==W1)
{
  needle_plot(V2, "V2", limits=range(ecg))
  needle_plot(W2, "W2", limits=range(ecg))
  needle_plot(W1, "W1", limits=range(ecg))
  graph_without_axis(ecg, "Original\nsignal", limits=range(ecg))
}
```

```{r wavelets-33}
c(V2[  1], g %*% V1[1:4])
c(V2[  2], g %*% V1[3:6])
c(V2[  3], g %*% V1[5:8])
c(V2[511], g %*% V1[1021:1024])
c(V2[512], g %*% V1[c(1023:1024, 1:2)])

@@

```{r wavelets-34}
quality_check <- mdwt(V1, daubcqf(4)$h.0, L=1)
quality_check_V2 <- quality_check$y[  1: 512]
quality_check_W2 <- quality_check$y[513:1024]
all(quality_check_V2==V2)
all(quality_check_W2==W2)
```

@@

```{r wavelets-35}
wt.3 <- mdwt(ecg, daubcqf(4)$h.0, L=3)
V3 <- wt.3$y[1:256]
W3 <- wt.3$y[257:512]
W2_check3 <- wt.3$y[513:1024]
W1_check3 <- wt.3$y[1025:2048]
all(W2_check3==W2)
all(W1_check3==W1)
{
  needle_plot(V3, "V3", limits=range(ecg))
  needle_plot(W3, "W3", limits=range(ecg))
  needle_plot(W2, "W2", limits=range(ecg))
  needle_plot(W1, "W1", limits=range(ecg))
  graph_without_axis(ecg, "Original\nsignal", limits=range(ecg))
}
```

@@

```{r wavelets-36}
c(V3[  1], g %*% V2[1:4])
c(V3[  2], g %*% V2[3:6])
c(V3[  3], g %*% V2[5:8])
c(V3[255], g %*% V2[509:512])
c(V3[256], g %*% V2[c(511:512, 1:2)])

@@

```{r wavelets-37}
wt.9 <- mdwt(ecg, daubcqf(4)$h.0, L=9)
V9 <- wt.9$y[   1:   4]
W9 <- wt.9$y[   5:   8]
W8 <- wt.9$y[   9:  16]
W7 <- wt.9$y[  17:  32]
W6 <- wt.9$y[  33:  64]
W5 <- wt.9$y[  65: 128]
W4 <- wt.9$y[ 129: 256]
W3 <- wt.9$y[ 257: 512]
W2 <- wt.9$y[ 513:1024]
W1 <- wt.9$y[1025:2048]
{
  needle_plot(V9, "V9", limits=range(ecg))
  needle_plot(W9, "W9", limits=range(ecg))
  needle_plot(W8, "W8", limits=range(ecg))
  needle_plot(W7, "W7", limits=range(ecg))
  needle_plot(W6, "W6", limits=range(ecg))
  needle_plot(W5, "W5", limits=range(ecg))
  needle_plot(W4, "W4", limits=range(ecg))
  needle_plot(W3, "W3", limits=range(ecg))
  needle_plot(W2, "W2", limits=range(ecg))
  needle_plot(W1, "W1", limits=range(ecg))
  graph_without_axis(ecg, "Original\nsignal", limits=range(ecg))
}
```

@@

```{r wavelets-38}
V9_diluted <- rep(0, 8)
V9_diluted[(1:4)*2-1] <- V9
W9_diluted <- rep(0, 8)
W9_diluted[(1:4)*2-1] <- W9

@@

```{r wavelets-39}
V8_diluted <- rep(0, 16)
V8_diluted[(1:8)*2-1] <- V8
W8_diluted <- rep(0, 16)
W8_diluted[(1:8)*2-1] <- W8

@@

```{r wavelets-40}
V7_diluted <- rep(0, 32)
V7_diluted[(1:16)*2-1] <- V7
W7_diluted <- rep(0, 32)
W7_diluted[(1:16)*2-1] <- W7

@@

```{r wavelets-41}

@@

```{r wavelets-42}

@@

```{r wavelets-43}
knitr::opts_chunk$set(fig.width=5, fig.height=5)
fn <- "https://www2.isye.gatech.edu/isyebayes/data/klaus256.asc"
img <- read.table(fn)
img <- 2*img-1
```

```{r wavelets-44}
img.plot <- function(imat,n=256) {
  xsq <- c(0,0,1,1)
  ysq <- c(0,1,1,0)
  par(mar=rep(0,4),xaxs="i",yaxs="i")
  plot(c(1,n+1),c(1,n+1),type="n",axes=FALSE)
  # imat <- imat/max(abs(imat))
  for (i in 1:n) {
    for (j in 1:n){
      d <- 0.5*imat[n+1-j,i]+0.5
      d <- min(1,max(0,d))
      polygon(xsq+i,ysq+j,density=-1,border=NA,col=rgb(d,d,d))
    }
  }
}

@@

```{r wavelets-45}
g <- daubcqf(8)$h.0
g
```

```{r wavelets-46}
h <- -daubcqf(8)$h.1
h
```

@@

```{r wavelets-47, fig.height=1}
d8_father_coefficients <- rep(0, 2048)
d8_father_coefficients[3] <- 1
d8_father <- as.vector(midwt(d8_father_coefficients, daubcqf(8)$h.0, L=8)$x)
d8_mother_coefficients <- rep(0, 2048)
d8_mother_coefficients[10] <- 1
d8_mother <- as.vector(midwt(d8_mother_coefficients, daubcqf(8)$h.0, L=8)$x)
{
  graph_without_axis(d8_father, "D8 father")
  graph_without_axis(d8_mother, "D8 mother")
}
```

@@

```{r wavelets-48}
wt.img.4 <- mdwt(img, daubcqf(8)$h.0, L=4)
reconstructed.img <- midwt(wt.img.4$y, daubcqf(8)$h.0, L=4)
img.plot(reconstructed.img$x)
```

```{r wavelets-49}
wt.img.1 <- mdwt(img, daubcqf(8)$h.0, L=1)
c(sum(img[1:8,1:8]*outer(g,g)), wt.img.1$y[ 1,   1])
c(sum(img[1:8,1:8]*outer(g,h)),wt.img.1$y[  1, 129])
c(sum(img[1:8,1:8]*outer(h,g)),wt.img.1$y[129,   1])
c(sum(img[1:8,1:8]*outer(h,h)),wt.img.1$y[129, 129])
```

@@

```{r wavelets-50}
x <- c(700, 300, 700)
y <- c(700, 300, 300)
m <- c(
  "W1\nh %*% g", 
  "W1\ng %*% h", 
  "W1\nh %*% h"
)
data.frame(x, y, m) %>%
  ggplot(aes(x, y, label=m)) +
  expand_limits(x=c(100, 900), y=c(100, 900)) + 
  geom_segment(x=100, y=100, xend=900, yend=100) +
  geom_segment(x=900, y=100, xend=900, yend=900) +
  geom_segment(x=900, y=900, xend=100, yend=900) +
  geom_segment(x=100, y=900, xend=100, yend=100) +
  geom_segment(x=500, y=100, xend=500, yend=900) +
  geom_segment(x=100, y=500, xend=900, yend=500) +
  theme_void() + 
  geom_text() -> layout_1
layout_1 +
    geom_text(x=300, y=700, label="V1\ng %*% g")
layout_1+
  geom_segment(x=300, y=500, xend=300, yend=900) +
  geom_segment(x=100, y=700, xend=500, yend=700) +
  geom_text(x=400, y=800, label="W2\nh %*% g") + 
  geom_text(x=200, y=600, label="W2\ng %*% h") + 
  geom_text(x=400, y=600, label="W2\nh %*% h") -> layout_2
layout_2 +
  geom_text(x=200, y=800, label="V2\ng %*% g")
layout_2 +
  geom_segment(x=200, y=700, xend=200, yend=900) +
  geom_segment(x=100, y=800, xend=300, yend=800) +
  geom_text(x=250, y=850, label="W3") + 
  geom_text(x=150, y=750, label="W3") + 
  geom_text(x=250, y=750, label="W3") + 
  geom_text(x=150, y=850, label="etc.")

@@

```{r wavelets-51}
gg1 <- matrix(0, nrow=256, ncol=256)
gg1[1:128, 1:128] <- wt.img.1$y[1:128, 1:128]
gg1_signal <- midwt(gg1, daubcqf(8)$h.0, L=1)  
img.plot(gg1_signal$x)
```

```{r wavelets-52}
gh1 <- matrix(0, nrow=256, ncol=256)
gh1[1:128, 129:256] <- wt.img.1$y[1:128, 129:256]
gh1_signal <- midwt(gh1, daubcqf(8)$h.0, L=1)  
img.plot(gh1_signal$x)
```

```{r wavelets-53}
hg1 <- matrix(0, nrow=256, ncol=256)
hg1[1:128, 129:256] <- wt.img.1$y[129:256, 1:128]
hg1_signal <- midwt(hg1, daubcqf(8)$h.0, L=1)  
img.plot(hg1_signal$x)
```

```{r wavelets-54}
hh1 <- matrix(0, nrow=256, ncol=256)
hh1[129:256, 129:256] <- wt.img.1$y[129:256, 129:256]
hh1_signal <- midwt(hh1, daubcqf(8)$h.1, L=1)  
img.plot(hh1_signal$x)
```

```{r wavelets-55}
gg4 <- matrix(0, nrow=256, ncol=256)
gg4[1:16, 1:16] <- wt.img.4$y[1:16, 1:16]
gg4_signal <- midwt(gg4, daubcqf(8)$h.0, L=4)  
img.plot(gg4_signal$x)
```
